Constant-round Blind Classical Verification of Quantum Sampling

### 2022-03-16 acknowledgements.tex
Do more cleanup

\section*{Acknowledgments}

The authors would like to thank Tomoyuki Morimae for his valuable feedback that helped improve the paper and for pointing out the related works \cite{takeuchi2018verification, hayashi2015verifiable}.
We are also thankful to anonymous reviewers for various useful comments. \Ethan{check if this is ok}

>>>>

\small{\subsubsection{Acknowledgments}

The authors would like to thank Tomoyuki Morimae for his valuable feedback that helped improve the paper and for pointing out the related works \cite{takeuchi2018verification, hayashi2015verifiable}.
We are also thankful to anonymous reviewers for various useful comments.

Kai-Min Chung is partially supported by the 2019 Academia Sinica Career Development Award under Grant no. 23-17, and MOST QC project under Grant no. MOST 108-2627-E-002-001.
This work was done while Yi Lee was affiliated to Academia Sinica and to National Taiwan University.
Part of this work was done while Han-Hsuan Lin was supported by Scott Aaronson's Vannevar Bush Faculty Fellowship from the US Department of Defense. Partially funded by MOST Grant no. 110-2222-E-007-002-MY3.
Xiaodi Wu is partially supported by the U.S. National Science Foundation grant CCF-1755800, CCF-1816695, and CCF-1942837 (CAREER).
}

### 2022-03-16 blind.tex

To resolve the issue, we let the verifier switch to a fresh new key for each round of the protocol. For example, when the prover $\Pblind$ returns the ciphertext of his first message, the verifier $\Vblind$ decrypts the ciphertext, computes his next message (in the clear), and then encrypt it using a fresh key $pk'$ and sends it to $\Pblind$.
Note that a fresh key pair is necessary here to ensure blindness, as decrypting uses information from the secret key.
Since the verifier $\Vblind$ only  sends fresh ciphertexts to $\Pblind$, this avoids the issue of circuit privacy. Additionally, to allow $\Pblind$ to homomorphically evaluate its next message, $\Vblind$ needs to encrypt the previous secret key $sk$ under the new public key $pk'$ and send it along with $pk'$ to $\Pblind$. This allows the prover to homomorphically convert ciphertexts under key $pk$ to ciphertexts under key $pk'$. By doing so, we show that for any cheating prover $\Pblindstar$, the interaction $(\Pblindstar, \Vblind)$ indeed simulates a valid interaction of $(\Pstar, V)$ for some cheating $\Pstar$, and hence the soundness of $\Pi$ implies the soundness of the compiled protocol. Finally, for the issue of the prover sending invalid ciphertexts, we note that this is not an issue if the decryption never fails, which can be achieved by simply let the decryption algorithm output a default dummy message (e.g., $0$) when it fails.

@@

Finally, we show the blindness of $\Piblind$ through a standard hybrid argument where the $sk_i$'s are ``erased" one by one, starting from $sk_T$.
Once $sk_1$ is eventually erased, $\QEnc(pk_1, x)$ and $\QEnc(pk_1, 0)$ become indistinguishable due to the IND-CPA security of $\QEnc$, and we complete the proof.
We now fill in the mathmatical details of this argument.

>>>>

To resolve the issue, we let the verifier switch to a fresh new key for each round of the protocol.\footnote{
An alternative strategy is to assume circuit privacy of $\QHE$. This seems to require many additional properties such as \emph{malicious} circuit privacy with efficient simulation and extraction when $\QGen$ is honest and secret key is available, multi-hop evaluation, and classical $\QEval$ on classical ciphertexts and circuits. While existing constructions such as \cite{circuit-private-qfhe} achieves some of these properties, we are unsure if any construction satisfies all of these requirements.}
For example, when the prover $\Pblind$ returns the ciphertext of his first message, the verifier $\Vblind$ decrypts the ciphertext, computes his next message (in the clear), and then encrypt it using a fresh key $pk'$ and sends it to $\Pblind$.
Note that a fresh key pair is necessary here to ensure blindness, as decrypting uses information from the secret key.
Since the verifier $\Vblind$ only  sends fresh ciphertexts to $\Pblind$, this avoids the issue of circuit privacy. Additionally, to allow $\Pblind$ to homomorphically evaluate its next message, $\Vblind$ needs to encrypt the previous secret key $sk$ under the new public key $pk'$ and send it along with $pk'$ to $\Pblind$. This allows the prover to homomorphically convert ciphertexts under key $pk$ to ciphertexts under key $pk'$. By doing so, we show that for any cheating prover $\Pblindstar$, the interaction $(\Pblindstar, \Vblind)$ indeed simulates a valid interaction of $(\Pstar, V)$ for some cheating $\Pstar$, and hence the soundness of $\Pi$ implies the soundness of the compiled protocol. Finally, for the issue of the prover sending invalid ciphertexts, we note that this is not an issue if the decryption never fails, which can be achieved by simply let the decryption algorithm output a default dummy message (e.g., $0$) when it fails.

@@

Finally, we show the blindness of $\Piblind$ through a standard hybrid argument where the $sk_i$'s are ``erased" one by one, starting from $sk_T$.
Once $sk_1$ is eventually erased, $\QEnc(pk_1, x)$ and $\QEnc(pk_1, 0)$ become indistinguishable due to the IND-CPA security of $\QEnc$.
We now fill in the details.

### 2022-03-16 hamiltonian_short.tex
Sync from Overleaf
Do more clean-up

We present the formal construction of our $\QPIP_1$ protocol $\PiSamp$ for $\SampBQP$ in \Cref{sec:sampbqp}, \Cref{ProtoQPIP1}. The soundness and completeness of \Cref{ProtoQPIP1} is stated in the following theorem, whose proof is also deferred to \Cref{sec:sampbqp}.% The proof of

@@

>>>>

We present the formal construction of our $\QPIP_1$ protocol $\PiSamp$ for $\SampBQP$ in our full version \cite{full-version}. The soundness and completeness of \Cref{ProtoQPIP1} is stated in the following theorem, whose proof is also deferred to \cite{full-version}.% The proof of

@@

\vspace{-3pt}

### 2022-03-16 introduction.tex

\paragraph{Overview of Mahadev's Protocol.}
%
Following~\cite{FOCS:Mahadev18a}, we define $\QPIP_{\tau}$ as classes of interactive proof systems between an (almost) classical verifier and a quantum prover, where the classical verifier has limited quantum computational capability, formalized as possessing $\tau$-qubit quantum memory. %\footnote{Intuitively, the size of local quantum memory limits the size of qubits that a general (entangled) quantum operation can be operated on at the same time. However, the verifier is still able to operate on all the qubits in a streaming fashion: i.e., receive the qubits sent from the prover and discard some local qubits in the case of overflow.}. 
A formal definition is given in Appendix~\ref{sec:qpip_def}.

@@

Another desirable property of CVQC protocols is  \emph{blindness}, which means that the prover does not learn any information about the private input for the delegated computation.\footnote{In literature, the definition of blindness may also require to additionally hide the computation. We note the two notions are equivalent from a feasibility point of view by a standard transformation (see Remark~\ref{rmk:blind-comp} in Section~\ref{sec:qpip_def}). 
%The two notions are equivalent from a feasibility point of view by a standard transformation that takes the description of the computation as part of the input and delegate the computation of  universal quantum circuits.
} In the relaxed setting where the verifier has a limited quantum capability, Hayashi and Morimae~\cite{hayashi2015verifiable} constructed a blind $\QPIP_1$ protocol for delegating quantum computation with information-theoretic security that also handles sampling problems. 
%
However, for purely classical verifiers, blind CVQC protocols seem much more difficult to construct. This goal is recently achieved by the seminal work of Gheorghiu and Vidick~\cite{FOCS:GheVid19}, who constructed the first blind CVQC protocol for $\BQP$ by constructing a composable remote state preparation protocol and combining it with the verifiable blind quantum computation protocol of Fitzsimons and Kashefi~\cite{FK17}. However, their protocol has   polynomially many rounds  and requires a rather involved analysis. Before our work, it is an open question whether constant-round blind CVQC protocol for $\BQP$ is achievable.

@@

%\vspace{2mm} \noindent \textbf{Organization.} \KM{check if its correct...} \KM{can use here to mention that preliminary is moved to appendix} 
\paragraph{Organization}
We provide preliminaries on technical background in Section~\ref{sec:prelim}. 
Our simulation-based definition of CVQC for $\SampBQP$ is discussed in Section~\ref{sec:samp_definition}. 
Our main technical contributions are explained in Section~\ref{sec:sampbqp_short} (a construction of $\QPIP_1$ protocol for $\SampBQP$), 
Section~\ref{sec:qpip0_all} (the construction of $\QPIP_0$ protocol for $\SampBQP$ based on the above $\QPIP_1$ protocol), 
and Section~\ref{sec:BlindBQP2} (a generic compiler to upgrade $\QPIP_0$ protocols with blindness).

>>>>

\paragraph{Overview of Mahadev's Protocol.}
%
Following~\cite{FOCS:Mahadev18a}, we define $\QPIP_{\tau}$ as classes of interactive proof systems between an (almost) classical verifier and a quantum prover, where the classical verifier has limited quantum computational capability, formalized as possessing $\tau$-qubit quantum memory. %\footnote{Intuitively, the size of local quantum memory limits the size of qubits that a general (entangled) quantum operation can be operated on at the same time. However, the verifier is still able to operate on all the qubits in a streaming fashion: i.e., receive the qubits sent from the prover and discard some local qubits in the case of overflow.}. 
A formal definition is given in our full version \cite{full-version}.

@@

Another desirable property of CVQC protocols is  \emph{blindness}, which means that the prover does not learn any information about the private input for the delegated computation.\footnote{In literature, the definition of blindness may also require to additionally hide the computation. We note the two notions are equivalent from a feasibility point of view by a standard transformation (see our full version \cite{full-version}). 
%The two notions are equivalent from a feasibility point of view by a standard transformation that takes the description of the computation as part of the input and delegate the computation of  universal quantum circuits.
} In the relaxed setting where the verifier has a limited quantum capability, Hayashi and Morimae~\cite{hayashi2015verifiable} constructed a blind $\QPIP_1$ protocol for delegating quantum computation with information-theoretic security that also handles sampling problems.
%
However, for purely classical verifiers, blind CVQC protocols seem much more difficult to construct. This goal is recently achieved by the seminal work of Gheorghiu and Vidick~\cite{FOCS:GheVid19}, who constructed the first blind CVQC protocol for $\BQP$ by constructing a composable remote state preparation protocol and combining it with the verifiable blind quantum computation protocol of Fitzsimons and Kashefi~\cite{FK17}. However, their protocol has   polynomially many rounds  and requires a rather involved analysis. Before our work, it is an open question whether constant-round blind CVQC protocol for $\BQP$ is achievable.

@@

%\vspace{2mm} \noindent \textbf{Organization.} \KM{check if its correct...} \KM{can use here to mention that preliminary is moved to appendix} 
\paragraph{Organization}
For preliminary technical background, see our full version \cite{full-version}. 
Our simulation-based definition of CVQC for $\SampBQP$ is discussed in Section~\ref{sec:samp_definition}. 
Our main technical contributions are explained in Section~\ref{sec:sampbqp_short} (a construction of $\QPIP_1$ protocol for $\SampBQP$), 
Section~\ref{sec:qpip0_all} (the construction of $\QPIP_0$ protocol for $\SampBQP$ based on the above $\QPIP_1$ protocol), 
and Section~\ref{sec:BlindBQP2} (a generic compiler to upgrade $\QPIP_0$ protocols with blindness).

### 2022-03-16 model.tex
Do more clean-up
Re-add accidentally deleted line

We next define what it means for a $\QPIP_\tau$ protocol\footnote{See Appendix~\ref{sec:qpip_def} for a formal definition of $\QPIP_\tau$.} to solve a $\SampBQP$ problem $\left(D_x\right)_{x\in\set{0, 1}^*}$.
Since sampling problems come with an accuracy parameter $\eps$, we let the prover $P$ and the verifier $V$ receive the input $x$ and $1^{1/\eps}$ as common inputs. 
Completeness is straightforward to define, which requires that when the prover $P$ is honest, the verifier $V$ should accept with high probability and output a sample $z$ distributed close to $D_x$ on input $x$. Defining soundness is more subtle. Intuitively, it requires that the verifier $V$ should never be ``cheated'' to accept and output an incorrect sample even when interacting with a malicious prover. We formalize this by a strong simulation-based definition, where we require that the joint distribution of the decision bit $d \in \set{\Acc, \Rej}$ and the output $z$ (which is $\bot$ when $d = \Rej$) is $\eps$-close (in either statistical or computational sense) to an ``ideal distribution'' $(d,z_{ideal})$, where $z_{ideal}$ is sampled from $D_x$ when $d = \Acc$ and set to $\bot$ when $d = \Rej$. Since the protocol receives the accuracy parameter $1^{1/\eps}$ as input to specify the allowed error, we do not need to introduce an additional soundness error parameter in the definition.

>>>>

We next define what it means for a $\QPIP_\tau$ protocol\footnote{See our full version \cite{full-version} for a formal definition of $\QPIP_\tau$.} to solve a $\SampBQP$ problem $\left(D_x\right)_{x\in\set{0, 1}^*}$.
Since sampling problems come with an accuracy parameter $\eps$, we let the prover $P$ and the verifier $V$ receive the input $x$ and $1^{1/\eps}$ as common inputs. 
Completeness is straightforward to define, which requires that when the prover $P$ is honest, the verifier $V$ should accept with high probability and output a sample $z$ distributed close to $D_x$ on input $x$. Defining soundness is more subtle. Intuitively, it requires that the verifier $V$ should never be ``cheated'' to accept and output an incorrect sample even when interacting with a malicious prover. We formalize this by a strong simulation-based definition, where we require that the joint distribution of the decision bit $d \in \set{\Acc, \Rej}$ and the output $z$ (which is $\bot$ when $d = \Rej$) is $\eps$-close (in either statistical or computational sense) to an ``ideal distribution'' $(d,z_{ideal})$, where $z_{ideal}$ is sampled from $D_x$ when $d = \Acc$ and set to $\bot$ when $d = \Rej$. Since the protocol receives the accuracy parameter $1^{1/\eps}$ as input to specify the allowed error, we do not need to introduce an additional soundness error parameter in the definition.
